---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mxd

<!-- badges: start -->

<!-- badges: end -->

The goal of mxd is to provide an *R* package for MaxDiff (i.e., Maximum Difference Scaling, also known as Best-Worst-Scaling case 1).

Currently, mxd provides the following function:

**Converting design to design matrix**

-   `bibd_to_dm()`: Function to convert balanced incomplete block design to design matrix used for estimation in mxd

-   `csv_to_dm()`: Function to convert .csv design from Lighthouse Studio to design matrix used for estimation in mxd

**Preparing design matrix for estimation in Stan**

-   `dm_to_stan_hb()`: Function to prepare input for HB estimation in Stan

-   `dm_to_stan_hb_cv()`: Function to prepare input for k-folded cross validation HB estimation in Stan

-   `dm_to_stan_mnl()`: Function to prepare input for MNL estimation in Stan

**Running estimation**

-   `mxd_hb()`: Function to run hierarchical Bayes estimation

-   `mxd_hb_cv()`: Function to estimate k-folded MaxDiff data

-   `mxd_logit()`: Function to run Bayesian version of MNL

**Checking convergence**

-   `convergence_stats()`: Function to test convergence stats of population's mean

Besides that, users can also use `shinystan()` to inspect convergence.

**Analyses tools**

-   `alphas_acf_plot()`: Function get posterior summary statistics for population's mean

-   `alphas()`: Function to get output of population's mean.

-   `beta_summary()`: Function to prepare posterior individual coefficients

-   `beta_post()`: Function to get posterior summary statistics for beta

-   `betas_violin()`: Function to get violin plot of posterior summary statistics

-   `bpe()`: Function to get the beta point estimates (i.e., aggregated individual coefficients)

-   `mxd_count()`: Function to run classical count analysis

-   `sigma_summary()`: Function to get posterior summary statistics for sigma

**Validation tools**

-   `post_hit()`: Function to calculate posterior hit rates (in-sample only)

-   `post_mhp()`: Function to calculate posterior mean hit probability (in-sample only)

-   `post_mae()`: Function to calculate posterior mean absolute error (both in-sample and out-of-sample)

-   `post_mae_cv()`: Function to calculate cross-validation mean absolute error

-   `post_medae()`: Function to calculate posterior median absolute error (both in-sample and out-of-sample)

-   `post_medae_cv()`: Function to calculate cross-validation median absolute error

-   `post_rmse()`: Function to calculate posterior root mean square error (both in-sample and out-of-sample)

-   `post_rmse_cv()`: Function to calculate cross-validation root mean square error

## Installation

You can install the development version of mxd from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("JoshSchramm94/mxd")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(mxd)
## basic example code
```
