---
title: "02 - mxd Bayesian MNL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{02 - mxd Bayesian MNL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses the Bayesian multinomial logit analysis.

# Setup

First, we load the relevant packages.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval = FALSE}
library(mxd)
```

# Create design matrix and input for stan

We begin by creating the design matrix using the `unanchored1` data frame that is part of the mxd package. In this data set, we simulated 100 participants, each completing 16 MaxDiff tasks with 4 alternatives per task. We start by creating the design matrix using the `csv_to_dm()` function.

```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::unanchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "best-worst"
)
```

Next, we prepare the input required for the `mxd_logit()` function. The mxd package relies on rstan [@rstan] for the estimation for the Bayesian multinomial logit analysis. rstan requires the input in a specific form, which can be created using the `dm_to_stan_mnl()` function. We specify:

-   the design matrix (`design`) created earlier

-   the participants' identifier (`id`)

-   the choice set indicator (`cs`)

-   the predictor variables (`items`)

-   the choice variable (`ch`)

-   the coding type used (`type`).

Finally, we need a prior for the beta coefficients which are drawn from a normal distribution. For more information, please check the [stan recommendations](https://github.com/stan-dev/stan/wiki/prior-choice-recommendations). Here, we choose a weakly informative prior.

```{r, eval = FALSE}
dm_stan <- dm_to_stan_mnl(
  design = design,
  id = id,
  cs = cs,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 10
)
```

# Estimation

Next, we can run the model. Here, we provide the `dm_stan` object created above to the `data_stan` argument and configure the model as follows:

-   `chains`: number of independent **M**onte **C**arlo **M**arkov **C**hains that should be run

-   `iter`: total number of iterations per chain (including warm-up)

-   `warmup`: number of warm-up iterations

-   `bw_size`: number of alternatives displayed in each MaxDiff task

-   `labels` (optional): to label the output

-   `seed`: set a seed for reproducibility

```{r, eval = FALSE}
out <- mxd_logit(
  data_stan = dm_stan,
  chains = 5L,
  iter = 2000,
  warmup = 1000,
  bw_size = 4,
  labels = c(paste0("item", seq.int(16))),
  seed = 12345
)
```

The output is a list of length 5 including the raw beta utilities, the zero-centered betas (range of 100), the choice probabilities, a summary, and finally the stan object.

# MaxDiff Coding

Alternatively, we display the coding for a direct anchored MaxDiff using the *maximum difference coding*. For demonstration purposes, we use the `anchored1` data frame from the mxd package.

First, we create the design matrix again using the `csv_to_dm()` function.

```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::anchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "maxdiff",
  anchor = "direct"
)
```

Next, we create the object needed for `mxd_logit()`. Here, we now change the `items` argument and also include the *anchor* (`item_17`). In addition, which is required only when using type `maxdiff` with a direct anchor, we also have to define the choice task of the anchor (i.e., in our case 17).

```{r, eval = FALSE}
dm_stan <- dm_to_stan_mnl(
  design = design,
  id = id,
  cs = cs,
  items = c(item_1:item_17),
  ch = choice,
  type = "maxdiff",
  prior_b = 10,
  anchor_start = 17
)
```

Finally, we can run our MNL model again.

```{r, eval = FALSE}
out <- mxd_logit(
  data_stan = dm_stan,
  chains = 5L,
  iter = 2000,
  warmup = 1000,
  bw_size = 4,
  labels = c(paste0("item", seq.int(17))),
  seed = 12345
)
```

# References
