---
title: "04 - mxd Posterior Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{04 - mxd Posterior Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial should teach how to analyse the posterior draws of the hierarchical Bayesian estimation.

First, we rerun the HB estimation.

# Setup

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval = FALSE}
library(mxd)
```


```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::unanchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "best-worst"
)

dm_stan <- dm_to_stan_hb(
  design = design,
  id = id,
  cs = cs,
  alt = alt,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 5,
  prior_omega = 2,
  prior_sigma = 2,
  demos = NULL
)

out <- mxd_hb(
  data_stan = dm_stan,
  chains = 5,
  iter = 2000,
  warmup = 1000,
  seed = 1910
)
```

# Analysis

## Hyperpriors b and sigma

First, we will focus on the hyperpriors of the model, specifically on the mean of the population-level parameters (`b`; see, e.g., @sablotny-wackershauser2024) and sigma. In the initial step, we check the convergence statistics using the `convergence_stats()` function, which provides the effective tail and bulk sample size as well as the Rhat values. We define the `stan_output` and also the parameters that we want to look at it. Here, we focus on the population-mean but instead we could also use sigma instead.

```{r, eval = FALSE}
convergence_stats(
  stan_output = out,
  pars = "b"
)
```

Next, we could plot the autocorrelation between the stored draws, using the `acf_plot()` function. Instead of defining `b` in the `pars` argument, we could also use `sigma`.

```{r, eval = FALSE}
acf_plot(
  stan_output = out,
  pars = "b"
)
```

Next, we run the `alphas()` function. Besides defining the object from `mxd_hb()` in `stan_output`, the number of alternatives shown per MaxDiff task (`bw_size`). In addition, we can optionally provide `labels` for the output and we have to define whether we estimated an anchored MaxDiff (`anchor`). The output of the `alphas()` function is a list of the length 4. It provides the stored alpha draws (`raw`) as well as the prepared one (zero-centered and probability) and finally a summary of all these statistics.

```{r, eval=FALSE}
alphas(
  stan_output = out,
  bw_size = 4,
  labels = paste0("item", seq.int(16)),
  anchor = FALSE
)
```

To get the summary statistics of sigma, we can run the `sigma_summary()` function.

```{r, eval = FALSE}
sigma_summary(
  stan_output = out
)
```

## Individual draws

Next, we focus on the individual coefficients (`beta`). In the initial step, we have to prepare them using the `betas_post()` function. Here we first define the `stan_output` argument. Afterwards, we define the number of alternatives shown per MaxDiff task (i.e., `bw_size`). To speed things up, we allow users to run the code using multiple cores. The default for `cores` is 1, however, this can be increased depending on the number of cores available. Finally, we again have to state whether we are using anchored MaxDiff (`anchor`) and optionally we can define labels.

```{r, eval = FALSE}
betas <- betas_post(
  stan_output = out,
  bw_size = 4,
  cores = 3,
  anchor = FALSE,
  labels = paste0("it", seq.int(16))
)
```

Once we prepared them, we can now access the betas. For example, if we want to summarize the raw beta draws, we can do so by using the `betas_summary()` function. In the `betas` argument, we need to define which betas we want to summarize. We can choose between the raw ones (`beta_raw`), the zero-centered interval scores (`betas_zc`), or the probability scores (`betas_prob`). For a better understanding of these scores, we refer to the documentation of Lighthouse Studio. Next, we need to define the variables (`vars`), and finally the participant's id (`id`). 

```{r, eval = FALSE}
betas_summary(
  betas = betas[["beta_raw"]],
  vars = c(it1:it16),
  id = id
)
```

To get the aggregated results for each individual (i.e., beta point estimates), you can use the `beta_point_estimates()` function.

```{r, eval = FALSE}
beta_point_estimates(
  betas = betas[["beta_raw"]],
  vars = c(it1:it16),
  id = id
)
```

To get a feeling of how the individual draws are distributed, you can use the `betas_violin()` function. The output of this function is a `ggplot` object (this is the same for the `acf_plot()` function) and can be further adjusted.

```{r, eval = FALSE}
betas_violin(
  betas = betas[["beta_raw"]],
  vars = c(it1:it15)
)
```

# Validation

Finally, we can use the posterior draws for validation by checking how well our model predicts validation tasks (see, e.g., @schramm2024). To predict choices from a validation, we need data for actual choices. For demonstration purposes, we provide the data frame `choicedata` in the mxd package.

```{r, eval = FALSE}
data("choicedata")
```

`HOT2` in `choicedata` is a validation task where the participants could choose from 8 alternatives, namely, items 4, 5, 7, 10, 11, 13, 14, and 16 (item 16 is `ref` in our case).

First, we will test the in-sample fit by examining the hit rate (i.e., the number of correct predicted choices). Here we define the posterior draws again (`betas_post`; these **have** to be raw draws), the data frame with the actual choices (`hot_data`), the `id` (which needs to have the same name in both data frames)[^1], the alternatives in the validation task (`opts`), the actual choice in the validation task (`hot_choice`). Finally, one can decide to get the raw posterior hit rates (i.e., the hit rates for each stored draw; set `raw` to `TRUE`) or the aggregated ones. The `post_mhp()`, which calculates the posterior mean hit probabilities.

[^1]: `betas_post()` automatically names the ids `id` . Thus, you might have to rename the variable in `hot_data` first.

```{r, eval = FALSE}
post_hit(
  betas_post = betas[["beta_raw"]],
  hot_data = choicedata,
  id = id,
  opts = c(it4, it5, it7, it10, it11, it13, it14, it16),
  hot_choice = HOT2,
  raw = FALSE
)
```

Alternatively, mxd also calculates he posterior mean absolute error (`post_mae()`), the posterior median absolute error (`post_medae()`) or the posterior root mean square error (`post_rmse()`). These 3 functions also work for out-of-sample validation. We generate some fake data for the same validation task shown in `post_hit()`.

```{r, eval = FALSE}
set.seed(1)
fake_hot2_data <- data.frame(
  hot2 = sample(
    x = seq.int(8),
    size = 1000,
    replace = TRUE
  )
)
```

Next, we can use, for example, the `post_mae()` function to calculate the mean absolute error for the posterior draws. The arguments are the same as for the `post_hit()` function except that we do not need to provide an `id` argument.

```{r, eval = FALSE}
post_mae(
  betas_post = betas[["beta_raw"]],
  hot_data = fake_hot2_data,
  opts = c(it4, it5, it7, it10, it11, it13, it14, it16),
  hot_choice = hot2,
  raw = FALSE
)
```

# References
