---
title: "mxd - Bayesian logit analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{02 - mxd Bayesian logit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses the Bayesian multinomial logit analysis.

# Setup

First, we again load the relevant packages.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(mxd)
library(magrittr)
library(dplyr)
```

# Create design matrix and input for stan

First, we again create the design matrix. We use the data frame `unanchored1` that comes with the mxd package. In this data set, 100 participants were simulated, each evaluating 16 MaxDiff tasks with 4 alternatives each.

```{r}
data("unanchored1")
```

Next, we create the design matrix using the `csv_to_dm()` function.

```{r}
design <- csv_to_dm(
    design = unanchored1,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = 16L,
    type = "best-worst"
  )
```

Next, we create the relevant input that we need for running the `mxd_logit()` function. The mxd package relies on rstan [@rstan] for the estimation for the Bayesian multinomial logit analysis. rstan needs the input in a specific form, which can be created using the `dm_to_stan_mnl()` function. Here, we first define the `design`, which is the design matrix we created in the code chunk above, the participants' identifier (`id`), the choice task indicator (`cs`), the predictors (`items`), the choice variable (`ch`), the coding used in the design matrix (`type`), and finally we need a prior for the beta coefficients derived from a normal distribution. For more information, please check the [stan recommendations](https://github.com/stan-dev/stan/wiki/prior-choice-recommendations). Here, we choose a weakly informative prior.

```{r}
dm_stan <- dm_to_stan_mnl(
  design = design,
  id = id,
  cs = cs,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 10
)
```

# Estimation

Next, we can run the model.

```{r}
out <- mxd_logit(
  data_stan = dm_stan,
  chains = 5L,
  iter = 2000,
  warmup = 1000,
  bw_size = 4,
  labels = c(paste0("item", seq.int(15))),
  seed = 12345
)
```

