---
title: "01 - mxd Choice Processes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01 - mxd Choice Processes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses different choice processes along with their corresponding coding.

# Setup

First, let us load the packages required for this vignette.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(mxd)
library(magrittr)
library(dplyr)
library(tidyr)
```

# Coding processes

Currently, mxd supports 7 coding processes, namely:

-   `best-worst`: best-worst coding (also known as marginal coding, see Formula 8 in @aizaki2023). This is the coding process implemented in Lighthouse Studio [@lighthouse2020]. Participants pick both the best and the worst item from the same set.

-   `best-worst-seq`: The best-worst sequential coding, which assumes that participants pick the best item which is then removed from the choice set for the worst item (see Formula 9 in @aizaki2023).

-   `worst-best-seq`: The worst-best sequential coding, which assumes that participants pick the worst item which is then removed from the choice set for the best item.

-   `maxdiff`: The maximum difference model, which assumes that the best choice and worst choice have the biggest difference across all pairs of two (see Formula 7 in @aizaki2023).

-   `exploded`: The exploded paired-comparison (see @chrzan2019), which creates every possible paired comparison for which one can assume the result.

-   `best-only`: Alternatively, one can also just use the best choices from the MaxDiff study...

-   `worst-only`: ... or the worst choices.

# Design Matrix

In this example, we simulate 5 MaxDiff tasks for one participant, each showing 3 items. The study includes a total of 5 items.

```{r}
K <- 5 # number of items
J <- 3 # size of tasks
I <- 1 # simulating one participant
T <- 4 # simulating one task

set.seed(1910)
mxd_task <- data.frame(
  id = rep(I, times = T * J),
  cs = rep(seq.int(T), each = J)
) %>%
  dplyr::mutate(
    item = sample(K, J),
    choice = sample(c(-1, 0, 1), J), .by = cs
  )
```

For illustrative purposes we focus on the first task only.

```{r}
head(mxd_task, n = J)
```

Here, participant `r I` chose item `r mxd_task[mxd_task$cs == 1 & mxd_task$choice == 1, "item"]` as the best item and item `r mxd_task[mxd_task$cs == 1 & mxd_task$choice == -1, "item"]` as the worst item.

## Best-worst coding

In the *best-worst* coding, two choice sets (`cs`) are created, one for the best choice (shown items display a `1`; see `cs` equals `1`) and one for the worst choice (shown items display a `-1`; see `cs` equals `2`).

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-worst"
  ),
  n = 6
)
```

## Best-worst-sequential coding

In the *best-worst-sequential* coding, the best choice (`r mxd_task[mxd_task$cs == 1 & mxd_task$choice == 1, "item"]`) is removed from the choice set for the worst item.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-worst-seq"
  ),
  n = 5
)
```

## Worst-best-sequential coding

In the *worst-best-sequential* coding, the best choice (`r mxd_task[mxd_task$cs == 1 & mxd_task$choice == -1, "item"]`) is removed from the choice set for the best item.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "worst-best-seq"
  ),
  n = 5
)
```

## Maximum difference coding

In the *maxdiff* coding, all possible pairs of 2 items are considered, and the pair with the greatest difference is selected.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "maxdiff"
  ),
  n = 6
)
```

## Exploded paired-comparison

In the *exploded paired comparison* coding, each possible paired comparison is coded. Here we know that the best choice wins against both the worst choice and the not chosen item and the not chosen item wins against the worst choice.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "exploded"
  ),
  n = 6
)
```

## Best-choice

Finally, when type is set to `best-only`, only the best choice is considered ...

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-only"
  ),
  n = 3
)
```

## Worst-choice

... and the same goes for `worst-only`, but only the worst choice is considered.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "worst-only"
  ),
  n = 3
)
```

# mxd with BWS from support.BWS

Researchers conducting a MaxDiff (i.e., BWS case 1) often rely on the well-supported package support.BWS [@aizaki2023]. The mxd package has the `bibd_to_dm()` function to convert a BWS study from support.BWS into the relevant format required for mxd.

First, we create a design. This is a highly simplified example, for design creation, please refer to the @aizaki2023. We simulate a data set with 100 participants (`I`), each participant evaluates a total of 7 tasks (`T`), each displaying 4 alternatives (`J`). We include a total of 7 items (`K`).

```{r}
I <- 100
T <- 7
J <- 4
K <- 7
```

Now, we can simulate the design, for reproducibility, we set a seed.

```{r}
set.seed(1910)
design <- tidyr::crossing(
  cs = seq.int(K), alt = seq.int(J)
) %>% 
  dplyr::mutate(item = sample(seq.int(K), J), .by = cs) %>% 
  tidyr::pivot_wider(
    names_from = alt,
    names_prefix = "item",
    values_from = item
)
```

Next, we can simulate choices for the 7 tasks. In support.BWS the alternative chosen is stored in choices (i.e., maximum `J`).

```{r}
dataset <- tidyr::crossing(
  id = seq.int(I), cs = seq.int(K), ch = c("b", "w")
) %>% 
  dplyr::mutate(
    cs = paste0(ch, cs),
    ch = sample(J, 2), .by = c(id, cs)) %>% 
  tidyr::pivot_wider(
    names_from = cs,
    values_from = ch
  )
```

Now, we can create the design. We define the `design`, the `data`, the name of the participants' identifier (`id`), the columns for the best choices (`best_ch`) and worst choices (`worst_ch`), respectively. Finally, we determine the `type` for coding (see above).

```{r}
output <- bibd_to_dm(
  design = design,
  data = dataset,
  id = id,
  best_ch = c(b1, b2, b3, b4, b5, b6, b7),
  worst_ch = c(w1, w2, w3, w4, w5, w6, w7),
  type = "best-worst"
)
```

# References
