---
title: "mxd - Analysis of posterior draws"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{04 - mxd Posterior Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial should teach how to analyse the posterior draws of the hierarchical Bayesian estimation.

First, we rerun the HB estimation.

# Setup

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mxd)

data("unanchored1")

design <- csv_to_dm(
    design = unanchored1,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = 16L,
    type = "best-worst"
  )

dm_stan <- dm_to_stan_hb(
  design = design,
  id = id,
  cs = cs,
  alt = alt,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 5,
  prior_omega = 2,
  prior_sigma = 2,
  demos = NULL,
  seed = 1910
)

out <- mxd_hb(
  data_stan = dm_stan,
  chains = 5,
  iter = 2000,
  warmup = 1000
)
```

# Analysis

## Hyperpriors b and sigma

First, we will focus on the hyperpriors of the model, specifically on the mean of the population-level parameters (`b`; see, e.g., @sablotny-wackershauser2024) and sigma. In the initial step, we check the convergence statistics using the `convergence_stats()` function, which provides the effective tail and bulk sample size as well as the Rhat values. We define the `stan_output` and also the parameters that we want to look at it. Here, we focus on the population-mean but instead we could also use sigma instead.

```{r}
convergence_stats(
  stan_output = out,
  pars = "b"
)
```

Next, we could plot the autocorrelation between the stored draws, using the `acf_plot()` function. Instead of defining `b` in the `pars` argument, we could also use `sigma`. 

```{r}
acf_plot(
  stan_output = out,
  pars = "b"
)
```

Next, we run the `alphas()` function. Besides defining the object from `mxd_hb()` in `stan_output`, the number of alternatives shown per MaxDiff task (`bw_size`). In addition, we can optionally provide `labels` for the output and we have to define whether we estimated an anchored MaxDiff (`anchor`). The output of the `alphas()` function is a list of the length 4. It provides the stored alpha draws (`raw`) as well as the prepared one (zero-centered and probability) and finally a summary of all these statistics. 

```{r, eval=FALSE}
alphas(
  stan_output = out,
  bw_size = 4,
  labels = paste0("item", seq.int(15)),
  anchor = FALSE
)
```

To get the summary statistics of sigma, we can run the `sigma_summary()` function.

```{r}
sigma_summary(
  stan_output = out
)
```

## Individual draws

Next, we focus on the individual coefficients (`beta`). In the initial step, we have to prepare them using the `betas_post()` function. Here we first define the `stan_output` argument. Afterwards, we define the number of alternatives shown per MaxDiff task (i.e., `bw_size`). To speed things up, we allow users to run the code using multiple cores. The default for `cores` is 1, however, this can be increased depending on the number of cores available. Finally, we again have to state whether we are using anchored MaxDiff (`anchor`) and optionally we can define labels. 

```{r}
betas <- betas_post(
  stan_output = out,
  bw_size = 4,
  cores = 3,
  anchor = FALSE,
  labels = paste0("it", seq.int(15))
)
```

