---
title: "02 - mxd Bayesian MNL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{02 - mxd Bayesian MNL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses the Bayesian multinomial logit analysis.

# Setup

First, we again load the relevant packages.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval = FALSE}
library(mxd)
```

# Create design matrix and input for stan

First, we again create the design matrix. We use the data frame `unanchored1` that comes with the mxd package. In this data set, 100 participants were simulated, each evaluating 16 MaxDiff tasks with 4 alternatives each.

Next, we create the design matrix using the `csv_to_dm()` function.

```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::unanchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "best-worst"
)
```

Next, we create the relevant input that we need for running the `mxd_logit()` function. The mxd package relies on rstan [@rstan] for the estimation for the Bayesian multinomial logit analysis. rstan needs the input in a specific form, which can be created using the `dm_to_stan_mnl()` function. Here, we first define the `design`, which is the design matrix we created in the code chunk above, the participants' identifier (`id`), the choice task indicator (`cs`), the predictors (`items`), the choice variable (`ch`), the coding used in the design matrix (`type`), and finally we need a prior for the beta coefficients derived from a normal distribution. For more information, please check the [stan recommendations](https://github.com/stan-dev/stan/wiki/prior-choice-recommendations). Here, we choose a weakly informative prior.

```{r, eval = FALSE}
dm_stan <- dm_to_stan_mnl(
  design = design,
  id = id,
  cs = cs,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 10
)
```

# Estimation

Next, we can run the model. Here, we provide the `dm_stan` object created above to the `data_stan` argument. Next, we define the number of independent **M**onte **C**arlo **M**arkov **C**hains that should be run in the `chains` argument. `iter` displays the number of iterations that should be run per chain and how many of those should be used for warm-up (`warmup`). In the `bw_size` argument, define the number of alternatives each MaxDiff task displayed. The `labels` argument is optional to give the output labels. The length of labels has to be the number of labels - 1 (the last item is used as reference level). Finally, we define a `seed` for reproducibility.

```{r, eval = FALSE}
out <- mxd_logit(
  data_stan = dm_stan,
  chains = 5L,
  iter = 2000,
  warmup = 1000,
  bw_size = 4,
  labels = c(paste0("item", seq.int(16))),
  seed = 12345
)
```

The output is a list of length 5 including the raw beta utilities, the zero-centered betas (range of 100), the choice probabilities, a summary statistic, and finally the stan object. 

# MaxDiff Coding

Alternatively, we display the coding for a direct anchored MaxDiff using the *maximum difference coding*. We use the `anchored1` data frame from the mxd package. 

First, we create the design matrix again again using the `csv_to_dm()` function.

```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::anchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "maxdiff",
  anchor = "direct"
)
```

And create the object that is needed for `mxd_logit()`. Here we now change the `items` argument and also include the *anchor* (`item_17`). In addition, which is unique to the type `maxdiff`, we also have to define the choice task of the anchor (i.e., in our case 17).

```{r, eval = FALSE}
dm_stan <- dm_to_stan_mnl(
  design = design,
  id = id,
  cs = cs,
  items = c(item_1:item_17),
  ch = choice,
  type = "maxdiff",
  prior_b = 10,
  anchor_start = 17
)
```

Finally, we can run our MNL model again.

```{r, eval = FALSE}
out <- mxd_logit(
  data_stan = dm_stan,
  chains = 5L,
  iter = 2000,
  warmup = 1000,
  bw_size = 4,
  labels = c(paste0("item", seq.int(17))),
  seed = 12345
)
```

# References
