---
title: "04 - mxd Posterior Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{04 - mxd Posterior Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial explains how to analyze the posterior draws from the Hierarchical Bayesian (HB) estimation.

First, we rerun the HB estimation.

# Setup

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval = FALSE}
library(mxd)

design <- csv_to_dm(
  design = mxd::unanchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "best-worst"
)

dm_stan <- dm_to_stan_hb(
  design = design,
  id = id,
  cs = cs,
  alt = alt,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 5,
  prior_omega = 2,
  prior_sigma = 2,
  demos = NULL
)

out <- mxd_hb(
  data_stan = dm_stan,
  chains = 5,
  iter = 2000,
  warmup = 1000,
  seed = 1910
)
```

# Analysis

## Hyperpriors b and sigma

First, we will focus on the hyperpriors of the model, specifically on the mean of the population-level parameters (`b`; see, e.g., @sablotny-wackershauser2024) and sigma. As a first step, we check the convergence statistics using the `convergence_stats()` function, which provides the effective sample size (tail and bulk) and the Rhat values. To run this function, we define the `stan_output` and also the parameters of interest. Here, we focus on the population means, but you could alternatively specify `sigma` in the `pars` argument.

```{r, eval = FALSE}
convergence_stats(
  stan_output = out,
  pars = "b"
)
```

Next, we can plot the autocorrelation between the stored draws using the `acf_plot()` function. As with the `convergence_stats()`function, you can specify either `b` or `sigma` in the `pars` argument.

```{r, eval = FALSE}
acf_plot(
  stan_output = out,
  pars = "b"
)
```

Next, we run the `b()` function. BWe need to specify the `stan_output` (estimated from `mxd_hb()`), and the number of alternatives shown per MaxDiff task (`bw_size`). Optionally, we can add `labels` for the output and we must indicate whether the design included anchored MaxDiff (`anchor`). The `alphas()` function returns a list of four elements: the raw alpha draws, zero-centered values, probability-scaled values, and a summary of these statistics.

```{r, eval=FALSE}
b(
  stan_output = out,
  bw_size = 4,
  labels = paste0("item", seq.int(16)),
  anchor = FALSE
)
```

To get the summary statistics of sigma, we can run the `sigma_summary()` function, which only requires the `stan_output`.

```{r, eval = FALSE}
sigma_summary(
  stan_output = out
)
```

## Individual draws

Next, we focus on the individual coefficients (`beta`). We begin by preparing them using the `betas_post()` function. First, we specify the `stan_output` argument, followed by the number of alternatives shown per MaxDiff task (i.e., `bw_size`). To speed up performance, the function supports parallel processing via the `cores` argument (default set to 1). Finally, we need to indication whether the design includes anchored MaxDiff (`anchor`) and optionally, we can define labels.

```{r, eval = FALSE}
betas <- betas_post(
  stan_output = out,
  bw_size = 4,
  cores = 3,
  anchor = FALSE,
  labels = paste0("it", seq.int(16))
)
```

Once the betas are prepared, we can access and analyze them. To summarize the raw beta draws, we use the `betas_summary()` function. In the `betas` argument, we specify which type of betas to summarize:

-   raw ones (`beta_raw`)

-   zero-centered interval scores (`betas_zc`)

-   probability scores (`betas_prob`)

For detailed explanations of these score types, please refer to the Lighthouse Studio documentation. Next, we need to define the variables (`vars`) as well as the participants' identifier (`id`). 

```{r, eval = FALSE}
betas_summary(
  betas = betas[["beta_raw"]],
  vars = c(it1:it16),
  id = id
)
```

To get the aggregated results for each individual (i.e., beta point estimates), you can use the `beta_point_estimates()` function.

```{r, eval = FALSE}
beta_point_estimates(
  betas = betas[["beta_raw"]],
  vars = c(it1:it16),
  id = id
)
```

To get an overview of how the individual draws are distributed, you can use the `betas_violin()` function. The output of this function is a `ggplot` object (this is the same for the `acf_plot()` function) and can be further adjusted.

```{r, eval = FALSE}
betas_violin(
  betas = betas[["beta_raw"]],
  vars = c(it1:it15)
)
```

# Validation

Finally, we can use the posterior draws to validate the model by assessing how well it predicts validation tasks (see, e.g., @schramm2024). To do this, we need actual choice data from a validation task. For demonstration purposes, we provide the data frame `choicedata` in the mxd package.

```{r, eval = FALSE}
data("choicedata")
```

`HOT2` in `choicedata` is a validation task where the participants could choose from 8 alternatives, namely, items 4, 5, 7, 10, 11, 13, 14, and 16.

First, we will test the in-sample fit by examining the hit rate (i.e., the number of correct predicted choices). Here, we define the posterior draws again (`betas_post`; these **have** to be raw draws), the data frame with the actual choices (`hot_data`), the `id` (which needs to have the same name in both data frames)[^1], the alternatives in the validation task (`opts`), the actual choice in the validation task (`hot_choice`). Finally, one can decide to get the raw posterior hit rates (i.e., the hit rates for each stored draw; set `raw` to `TRUE`) or the aggregated ones. The `post_mhp()`, which calculates the posterior mean hit probabilities, works identically to the `post_hit()` function.

[^1]: `betas_post()` automatically names the ids a column named `id` . Thus, you might have to rename the variable in `hot_data` first.

```{r, eval = FALSE}
post_hit(
  betas_post = betas[["beta_raw"]],
  hot_data = choicedata,
  id = id,
  opts = c(it4, it5, it7, it10, it11, it13, it14, it16),
  hot_choice = HOT2,
  raw = FALSE
)
```

Alternatively, the mxd packge provides additional validation metrics:

-   posterior mean absolute error (`post_mae()`)

-   posterior median absolute error (`post_medae()`)

-   posterior root mean square error (`post_rmse()`)

These 3 functions also work for out-of-sample validation. For demonstration purposes, we generate some fake data for the same validation task shown in `post_hit()`.

```{r, eval = FALSE}
set.seed(1)
fake_hot2_data <- data.frame(
  hot2 = sample(
    x = seq.int(8),
    size = 1000,
    replace = TRUE
  )
)
```

Next, we can use, for example, the `post_mae()` function to calculate the mean absolute error for the posterior draws. The arguments are the same as for the `post_hit()` function except that we do not need to define an `id` argument.

```{r, eval = FALSE}
post_mae(
  betas_post = betas[["beta_raw"]],
  hot_data = fake_hot2_data,
  opts = c(it4, it5, it7, it10, it11, it13, it14, it16),
  hot_choice = hot2,
  raw = FALSE
)
```

# References
