---
title: "01 - mxd Choice Processes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01 - mxd Choice Processes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses the different choice processes and the corresponding coding processes.

# Setup

First, let us load the packages required for this vignette.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(mxd)
library(magrittr)
library(dplyr)
library(tidyr)
```

# Coding processes

Currently, mxd provides a total of 7 different coding processes, namely:

-   `best-worst`: best-worst coding (also known as marginal coding, see formula 8 in @aizaki2023). This is the coding process implemented in Lighthouse Studio [@lighthouse2020]. Participants make the best and the worst choices from the same set.

-   `best-worst-seq`: The best-worst sequential coding, which assumes that participants first choose the best option and then the worst option from a set excluding the best option (see formula 9 in @aizaki2023).

-   `worst-best-seq`: The worst-best sequential coding, which assumes that participants first choose the worst option and then the best option from a set excluding the worst option.

-   `maxdiff`: The maximum difference model, which assumes that the best choice and worst choice have the biggest difference across all pairs of two (see formula 7 in @aizaki2023).

-   `exploded`: The exploded paired-comparison (see @chrzan2019), which creates every possible paired comparison for which one can assume the result.

-   `best-only`: Alternatively, one can also just use the best choices from the MaxDiff study...

-   `worst-only`: ... or the worst choices.

# Design Matrix

For this example, we will simulate 5 MaxDiff tasks for one participant showing 3 items per task. Further, we assume that a total of 5 items were included in this study.

```{r}
K <- 5 # no of items
J <- 3 # size of tasks
I <- 1 # simulating one participant
T <- 4 # simulating one task

set.seed(1910)
mxd_task <- data.frame(
  id = rep(I, times = T * J),
  cs = rep(seq.int(T), each = J)
) %>%
  dplyr::mutate(
    item = sample(K, J),
    choice = sample(c(-1, 0, 1), J), .by = cs
  )
```

In the following, we will focus on the first MaxDiff task only.

```{r}
head(mxd_task, n = J)
```

In this example, participant `r I` chose item `r mxd_task[mxd_task$cs == 1 & mxd_task$choice == 1, "item"]` as best item and item `r mxd_task[mxd_task$cs == 1 & mxd_task$choice == -1, "item"]` as worst item.

## Best-worst coding

In the *best-worst* coding, two choice sets (`cs`) are created, one for the best choice (shown items display a `1`; see `cs` equals `1`) and one for the worst choice (shown items display a `-1`; see `cs` equals `2`).

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-worst"
  ),
  n = 6
)
```

## Best-worst-sequential coding

In the *best-worst-sequential* coding, the best choice (`r mxd_task[mxd_task$cs == 1 & mxd_task$choice == 1, "item"]`) is removed from the choice set for the worst item.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-worst-seq"
  ),
  n = 5
)
```

## Worst-best-sequential coding

In the *worst-best-sequential* coding, the best choice (`r mxd_task[mxd_task$cs == 1 & mxd_task$choice == -1, "item"]`) is removed from the choice set for the best item.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "worst-best-seq"
  ),
  n = 5
)
```

## Maximum difference coding

In the *maxdiff* coding, each possible pair of two is coded and the one that has the maximum difference is chosen.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "maxdiff"
  ),
  n = 6
)
```

## Exploded paired-comparison

In the *exploded paired comparison* coding, each possible paired comparison is coded. Here we know that the best choice wins against both the worst choice and the not chosen item and the not chosen item wins against the worst choice.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "exploded"
  ),
  n = 6
)
```

## Best-choice

Finally, when type is set to `best-only`, only the best choice is considered ...

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "best-only"
  ),
  n = 3
)
```

## Worst-choice

... and the same goes for `worst-only`, but only the worst choice is considered.

```{r}
head(
  csv_to_dm(
    design = mxd_task,
    id = id,
    cs = cs,
    item = item,
    ch = choice,
    mxd_tasks = T,
    type = "worst-only"
  ),
  n = 3
)
```

# mxd with BWS from support.BWS

Researchers that conduct a MaxDiff (i.e., BWS case 1) often rely on the great package support.BWS [@support.BWS]. mxd has the function `bibd_to_dm()` to convert a BWS study from support.BWS to the relevant design required for mxd.

First, we create a design. This is a highly simplified example, for design creation, please refer to the @support.BWS package. We simulate a data set with 100 participants (`I`), each participant evaluates a total of 7 tasks (`T`), each displaying 4 alternatives (`J`). We include a total of 7 items (`K`).

```{r}
I <- 100
T <- 7
J <- 4
K <- 7
```

Now, we can simulate the design, for reproducibility, we set a seed.

```{r}
set.seed(1910)
design <- tidyr::crossing(
  cs = seq.int(K), alt = seq.int(J)
) %>% 
  dplyr::mutate(item = sample(seq.int(K), J), .by = cs) %>% 
  tidyr::pivot_wider(
    names_from = alt,
    names_prefix = "item",
    values_from = item
)
```

Next, we can simulate choices for the 7 tasks. In support.BWS the alternative chosen is stored in choices (i.e., maximum `J`).

```{r}
dataset <- tidyr::crossing(
  id = seq.int(I), cs = seq.int(K), ch = c("b", "w")
) %>% 
  dplyr::mutate(
    cs = paste0(ch, cs),
    ch = sample(J, 2), .by = c(id, cs)) %>% 
  tidyr::pivot_wider(
    names_from = cs,
    values_from = ch
  )
```

Now, we can create the design. We define the `design`, the `data`, the name of the participants' identifier (`id`), the columns for the best choices (`best_ch`) and worst choices (`worst_ch`), respectively. Finally, we determine the `type` for coding (see above)

```{r}
output <- bibd_to_dm(
  design = design,
  data = dataset,
  id = id,
  best_ch = c(b1, b2, b3, b4, b5, b6, b7),
  worst_ch = c(w1, w2, w3, w4, w5, w6, w7),
  type = "best-worst"
)
```

# References
