---
title: "03 - mxd Hierarchical Bayesian MNL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{03 - mxd Hierarchical Bayesian MNL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

This tutorial discusses the Hierarchical Bayesian (HB) multinomial logit analysis.

# Setup

First, we again load the relevant packages.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval = FALSE}
library(mxd)
```

# Preparation

The first preparation steps are similar to the one shown in the tutorial of the [Bayesian MNL analysis](mxd-logit.html), thus, we do not discuss them in detail. 

```{r, eval = FALSE}
design <- csv_to_dm(
  design = mxd::unanchored1,
  id = id,
  cs = cs,
  item = item,
  ch = choice,
  mxd_tasks = 16L,
  type = "best-worst"
)
```

Next, we prepare the input for the `mxd_hb()` function to estimate the utilities. To do so, we use the `dm_to_stan_hb()` function, which is quite similar to the `dm_to_stan_mnl()` function. For this function, we additionally have to define the `alt` argument, which is the variable that indicates the alternatives within a MaxDiff task. In addtion, we have to define additional priors, namely, for omega (`prior_omega`) and sigma (`prior_sigma`). Again, we define weakly informative priors (i.e., the default values of the function). Lastly, we also state that there are no *Z* variables (i.e., no demographics included; `demos` is set to `NULL`).

```{r, eval = FALSE}
dm_stan <- dm_to_stan_hb(
  design = design,
  id = id,
  cs = cs,
  alt = alt,
  items = c(item_1:item_16),
  ch = choice,
  type = "best-worst",
  prior_b = 5,
  prior_omega = 2,
  prior_sigma = 2,
  demos = NULL
)
```


# Estimation 

Finally, we can run the HB estimation by using the `mxd_hb()` function. We define the object (`dm_stan`) created in the chunk above in the `data_stan` argument, and again define the number of chains to be run (`chains`), the number of iterations to be run (`iter`) and how many of those should be used for warm-up purposes (`warmup`).

```{r, eval = FALSE}
out <- mxd_hb(
  data_stan = dm_stan,
  chains = 5,
  iter = 2000,
  warmup = 1000
)
```

