---
title: "mxd - an R package for MaxDiff in R"
tags:
- R
- MaxDiff
- Hierarchical Bayes
authors:
- name: Joshua Benjamin Schramm
  orcid: 0000-0001-5602-4632
  corresponding: true
  affiliation: 1
- name: Daniel Guhl
  orcid: 0000-0003-2114-9405
  corresponding: false
  affiliation: 2
- name: Marcel Lichters
  orcid: 0000-0002-3710-2292
  corresponding: false
  affiliation: 1
affiliations:
- name: Otto-von-Guericke-University Magdeburg
  index: 1
- name: Humboldt-Universität zu Berlin
  index: 2
date: May 2025
output: html_document
bibliography: paper.bib
---

# Introduction

xxx

# Motivation

xxx

# Statement of need

xxx

# Workflow

## Preparation

In the first step, we load the `mxd` package and will have a look at the design how it is usually exported.

```{r}
library(mxd)
design <- readRDS("../data/design.rds")
head(design)
```

To convert this into the right format, we use the `csv_to_dm()` function. For the `anchor` argument (default set to `NULL`), the user can choose from both `direct` or `indirect`. For the `type` of coding, the user can decide between `best-worst`, `best-worst-seq` (first best, afterwards worst from a reduced choice set), `worst-best-seq`, `best-only` (only considering best choices), `worst-only`, `maxdiff` (maximum difference), and `exploded` (exploded logit paired comparison). For our example, we stick to the industry standard.

```{r}
dm <- csv_to_dm(
  design = design,
  id = ID,
  cs = Set,
  item = Item,
  ch = Response,
  anchor = "direct",
  mxd_tasks = 16L,
  type = "best-worst"
)

head(dm)
```

Often, users relying on *R* use the `crossdes` [@crossdes] or `support.BWS` [@support.BWS] packages. To convert their created functions and data frames in the correct format, we implemented the `bibd_to_dm()` function. To show how to prepare the design matrix, when the design was created with the `crossdes` package, we provide the following example. To prepare the design and a fake data set, we use the exemplary code provided in `support.BWS` [@support.BWS].

```{r}
library(support.BWS)
library(crossdes)
set.seed(1)
bibd_design <- find.BIB(trt = 7, b = 7, k = 4)
items <- c("apple", "orange", "grapes", "banana", "peach", "melon", "pear")
b <- c(1, 0.6, 0.3, -0.3, -0.5, -1, 0)
dataset <- bws.response(design = bibd_design, item.names = items, b = b, n = 100, seed = 1)
```

Next, we can apply the `bibd_to_dm()` function provided by the `mxd` package

```{r}
bibd_dm <- bibd_to_dm(
  design = bibd_design,
  data = dataset,
  id = id,
  best_ch = c(B1, B2, B3, B4, B5, B6, B7),
  worst_ch = c(W1, W2, W3, W4, W5, W6, W7),
  type = "best-worst"
)
```


## Estimation

### Classical frequentist MNL

`mxd` provides the `mxd_logit()`, which relies on the `logitr` [@logitr] to run MNL. `bw_size` defines the size of the MaxDiff tasks. The output provides both raw utilities as well as centered one and important scores.

```{r}
mxd_logit(
  data = dm,
  ch = choice,
  cs = Set,
  items = c(Item_1:Item_16),
  bw_size = 4,
  reference = "anchor",
  anchor = TRUE
)
```

### Bayesian MNL

Alternatively, the user can also use a Bayesian approach to run the MNL model. `mxd` relies on `rstan` to run the estimation. The model code is the following:

```{r}
cat(rstan::get_stancode(mxd:::stanmodels$mnl))
```

First, we need to prepare the input for running the model, for which we can use the `dm_to_stan_mnl()` function. The default prior for `b` is set to `5` (i.e., weakly informative prior), however, can be adapted as shown in the example below.

```{r}
data_stan <- dm_to_stan_mnl(
  design = dm, 
  id = ID, 
  cs = Set, 
  alt = alt, 
  items = c(Item_1:Item_17), 
  ch = choice, 
  prior_b = 10)
```

Next, we can run the estimation. For demonstration purposes, we lower the number of iterations and just use 2 chains.

```{r, warning = FALSE}
output <- mxd_logit_bayesian(
  data_stan = data_stan,
  chains = 2L,
  iter = 400L,
  warmup = 200L,
  bw_size = 4L,
  labels = paste0("g", seq_len(16)),
  anchor = TRUE,
  seed = 1910L
)
```

Running this function provides

### Bayesian HB-MNL

Alternatively, we could also run a hierarchical Bayesian MNL model. The Stan code has the benefits that it allows for alternating choice set sizes, e.g., it can be implemented with each type of coding we introduced above.

```{r}
cat(rstan::get_stancode(mxd:::stanmodels$hbmnl))
```

To run the function, we first need to prepare the input accordingly. To do so, the users can apply the `dm_to_stan_hb()` function.

```{r}
stan_input <- dm_to_stan_hb(
    design = dm, 
    id = ID, 
    cs = Set,
    alt = alt,
    items = c(Item_1:Item_17), 
    ch = choice,
    prior_b = NULL,
    prior_omega = NULL,
    prior_sigma = NULL,
    demos = NULL
    )
```

Next, we can estimate the model.

```{r, warning = FALSE}
output <- mxd_hb(
  data_stan = stan_input[["stan_input"]],
  chains = 5L,
  iter = 400L,
  warmup = 200L,
  seed = 1910L,
)
```

### Count Analysis

For those users who are prefer to run classical count analysis, we implemented the `mxd_count()` function. The user has to provide an unanchored design to run the code. To do, we will adapt the `design` data frame briefly.

```{r}
unanc_design <- dplyr::filter(design, Set <= 16)
```

Afterwards, we can run `mxd_count()`

```{r}
mxd_count(
  data = unanc_design,
  cs = Set,
  item = Item,
  ch = Response,
  no_items = 16L,
  labels = paste0("Label ", seq_len(16))
)
```

## Convergence

We could also check the convergence after estimating the model. Of course, users could also use `shinystan` [@shinystan]. `mxd` provides, for example, the `acf_plot()` function to check autocorrelation between draws. `acf_plot()` exports a `ggplot2` object that can be further adjusted by the user. Either `sigma` or `b` can be plotted.

```{r}
acf_plot(
  stan_output = output,
  pars = "b",
  labels = paste0("G", seq_len(16))
)
```

Also, we could look directly at `b`'s Rhat values and effective sample sizes.

```{r}
convergence_stats(
  stan_output = output,
  pars = "b",
  labels = paste0("G", seq_len(16))
)
```

## Parameter summary

### `b`

To summarize the `b` draws, `mxd` provides the `alphas()` function.

```{r}
alpha_scores <- alphas(
  stan_output = output,
  bw_size = 4,
  labels = paste0("G", seq_len(16)),
  anchor = TRUE
)
```

### `sigma`

To summarize the `sigma` draws, `mxd` provides the `sigma_summary()` function.

```{r}
sigma_summary(
  stan_output = output,
  labels = paste0("G", seq_len(16))
)
```

### `beta`

`mxd` provides multiple functions for the `beta` scores (i.e., individual utilities). In the initial step, the user should prepare the draws accordingly, which is possible using the `betas_post()` function.

```{r}
betas_prep <- betas_post(
  stan_output = output,
  bw_size = 4, 
  cores = 4L,
  ids = stan_input[["ids"]][["orig_id"]],
  labels = paste0("g", seq_len(16)),
  anchor = TRUE
)
```

Once prepared, we can get the summary statistic of the results. Since `betas_prep()` provides three different types of `beta` preparations, namely, the raw, the zero-anchored, and the choice probabilities scores, we can choose which we want to summarise. In the example below, we will use the raw utilties.

```{r}
betas_summary(
  betas = betas_prep[["beta_raw"]],
  vars = c(g1:ref)
)
```

Alternatively, we could also show the choice probabilities.

```{r}
betas_summary(
  betas = betas_prep[["beta_prob"]],
  vars = c(g1:ref)
)
```

Often researchers use the aggregated utilities across all posterior draws. To get these, users can use the `bpe()` (i.e., beta point estimates) function.

```{r}
bpe(
  betas = betas_prep[["beta_raw"]],
  vars = c(g1:ref),
  id = id
) |>
  head()
```

To plot the beta scores, `mxd` provides the `beta_violin()` function, which, similar to the `acf_plot()` provides a `ggplot2` object as output and thus can we further adjusted by the user. For example below, we adjust the colors of the items.

```{r}
betas_violin(
  betas = betas_prep[["beta_raw"]],
  vars = c(g1:g16),
  labels = paste0("g", seq_len(16))
) + 
  ggplot2::scale_fill_manual(values = rev(c("red","blue", "green", "yellow",
                                   "orange", "magenta", "grey", "black",
                                   "pink", "darkgreen", "lightgreen", "brown",
                                   "lightblue", "darkblue", "lightgrey", "darkgrey")))
```

## Validation

Similar to the `validateHOT` package by [@schramm2025a], `mxd` also provides tools for validation. However, the main difference is that `validateHOT` is intended to work with the beta point estimates, while `mxd` works with the posterior draws and in addition, some of its functions can also be applied to a validation sample.

### Internal fit

First, we load an additional data frame, in which we stored the actual choice in the holdout task.

```{r}
hot_data <- read.csv("../data/data.csv")
```

For internal fit, `mxd` provides both the `post_hit()` function to calculate the hit rate across the posterior draws ...

```{r}
post_hit(
  betas_post = betas_prep[["beta_raw"]],
  hot_data = hot_data,
  id = id,
  opts = c(g1, g8, g9, g13, g14, g15, g16, ref),
  hot_choice = HOT1,
  raw = FALSE
)
```

and also the `post_mhp()` function to calcuate the mean hit probability across posterior draws

```{r}
post_mhp(
  betas_post = betas_prep[["beta_raw"]],
  hot_data = hot_data,
  id = id,
  opts = c(g1, g8, g9, g13, g14, g15, g16, ref),
  hot_choice = HOT1,
  raw = FALSE
)
```

### External fit

The functions `post_mae()`, `post_medae()`, and `post_rmse()` can be used for both internal as well as external fit. `post_mae()` calculates the mean absolute error,

```{r}
post_mae(
  betas_post = betas_prep[["beta_raw"]],
  hot_data = hot_data,
  opts = c(g1, g8, g9, g13, g14, g15, g16, ref),
  hot_choice = HOT1,
  raw = FALSE
)
```

`post_medae()` calculates the median absolute error,

```{r}
post_medae(
  betas_post = betas_prep[["beta_raw"]],
  hot_data = hot_data,
  opts = c(g1, g8, g9, g13, g14, g15, g16, ref),
  hot_choice = HOT1,
  raw = FALSE
)
```

and finally, `post_rmse()` calculates the root mean square error.

```{r}
post_rmse(
  betas_post = betas_prep[["beta_raw"]],
  hot_data = hot_data,
  opts = c(g1, g8, g9, g13, g14, g15, g16, ref),
  hot_choice = HOT1,
  raw = FALSE
)
```

## Cross-validation
